Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-04-09 paper.qmd
Add plots for 2D model construction

nldr_scurve <- umap_scurve |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
  interior_annotation("a")
```

@@

@@

@@

```{r}
#| label: fig-NLDR-scurve
#| echo: false
#| fig-cap: "Six different NLDR representations of the same data. Different techniques and different parameter choices are used. Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical parameter choice. Would they make different decisions downstream in the analysis depending on which version seen? Which is the most accurate representation of the structure in high dimensions?"
#| fig-width: 8
#| fig-height: 4
#| out-width: 100%

nldr_scurve +
  plot_layout(ncol = 4)
```

>>>>

umap_scurve_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_scurve, 
                                    x = "UMAP1", y = "UMAP2"))) |>
  rename(c("UMAP1" = "scaled_UMAP1", 
                  "UMAP2" = "scaled_UMAP2")) |>
  mutate(ID = 1:NROW(umap_scurve))

nldr_scurve <- umap_scurve_scaled |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
  interior_annotation("a")
```

@@

```{r}
num_bins_x_scurve <- 6
num_bins_y_scurve <- 14
hex_size_scurve <- 0.13

hb_obj_scurve <- hex_binning(data = umap_scurve_scaled, x = "UMAP1", 
                      y = "UMAP2", num_bins_x = num_bins_x_scurve, 
                      num_bins_y = num_bins_y_scurve, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_scurve, 
                      col_start = "UMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_scurve$centroids))

## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_scurve$hex_poly))

## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_scurve$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)

hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))

hex_grid_scurve <-  ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id), fill = "#ffffff") +
  geom_point(data = umap_scurve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  interior_annotation("b")
```

@@

```{r}
hex_grid_nonempty <- hex_grid |>
  filter(hex_poly_id %in% df_bin_centroids$hexID)

hex_grid_nonempty_scurve <-  ggplot(data = hex_grid_nonempty, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id), fill = "#ffffff") +
  geom_point(data = df_bin_centroids, aes(x = c_x, y = c_y), color = "#33a02c") +
  interior_annotation("c") 
```

@@

```{r}
## Triangulate bin centroids
tr1_object_scurve <- tri_bin_centroids(df_bin_centroids, x = "c_x", y = "c_y")
tr_from_to_df_scurve <- gen_edges(tri_object = tr1_object_scurve)

## Compute 2D distances
distance_scurve <- cal_2d_dist(tr_coord_df = tr_from_to_df_scurve, 
                             start_x = "x_from", start_y = "y_from", 
                             end_x = "x_to", end_y = "y_to", 
                             select_vars = c("from", "to", "distance"))

## To find the benchmark value
benchmark_scurve <- find_lg_benchmark(distance_edges = distance_scurve, distance_col = "distance")

trimesh_removed_scurve <- vis_rmlg_mesh(distance_edges = distance_scurve, 
                                        benchmark_value = benchmark_scurve, 
                                        tr_coord_df = tr_from_to_df_scurve, 
                                        distance_col = "distance") +
  interior_annotation("d") 
```

```{r}
#| label: fig-NLDR-scurve
#| echo: false
#| fig-cap: ""
#| fig-width: 8
#| fig-height: 4
#| out-width: 100%

nldr_scurve + hex_grid_scurve + hex_grid_nonempty_scurve + trimesh_removed_scurve +
  plot_layout(ncol = 4)
```
